---
title: "In-class Ex04"
author: "Michelle Zhang"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

### 1. Install and loading R packages

```{r}
pacman::p_load(plotly, DT, patchwork, ggstatsplot, readxl, performance, parameters, see, gtsummary, tidyverse)
```

### 2. Importing Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

```{r}
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
```

### 3. Interactivity in plotting

```{r}
plot_ly(data = exam_data,
        x = ~ENGLISH,
        y = ~MATHS,
        color = ~RACE)
```

```{r}
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH,
                color = RACE)) +
  geom_point(size = 1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p) 
```

### 4. Visual statistical plotting

#### 4.1 Two-sample mean testing

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER,
  y = MATHS,
  #"p" is parametric test while "np" is non-parametric test
  type = "p",
  messages = FALSE
)
```

#### 4.2 Scatterplot testing

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  #the default for marginal is TRUE which will show the marginal plots
  marginal = TRUE
)
```

### 5. Model visualization

#### 5.1 Building least-square multiple regression model

`lm()` is base R model to build least-square multiple regression model

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM +
              Weight + Guarantee_Period, data = car_resale)
model
```

#### 5.2 Diagnostic test : Check for multi-collinearity

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

#### 5.3 Diagnostic test : Check for normality assumption

```{r}
#Remove Mfg_Year from model due to high collinearity
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
model1
```

```{r}
check_n <- check_normality(model1)
plot(check_n)
```

#### 5.4 Diagnostic test : Check for variance homogeneity

```{r}
check_h <- check_heteroscedasticity(model1)
plot(check_h)
```

#### 5.5 Diagnostic test : Check for everything

```{r}
#| fig-height: 15
#| fig-width: 12.5
check_model(model1)
```

#### 5.6 Visualizing regression parameters

```{r}
plot(parameters(model1))
```

```{r}
ggcoefstats(model1, 
            output = "plot")
```

### 6. Visualization of uncertainty

Data preparation

```{r}
#group by RACE and calculate mean, sd, and se of MATHS score
my_sum <- exam_data |> 
  group_by(RACE) |> 
  summarize(
    n = n(),
    mean = mean(MATHS),
    sd = sd(MATHS)) |>
  mutate(se = sd/sqrt(n-1))
```

Plotting using ggplot2

```{r}
ggplot(my_sum) +
  
  geom_errorbar(
    aes(x = RACE,
        ymin = mean - se,
        ymax = mean + se),
    width = 0.2,
    colour = "black",
    alpha = 0.9,
    linewidth = 0.5) +
  
  geom_point(
    aes(x = RACE,
        y = mean),
    stat = "identity",
    colour = "red",
    size = 1.5,
    alpha = 1) +
  
  ggtitle("Standard error of mean
          maths score by race")
```
