---
title: "In-class Exercise 2"
author: "Michelle Zhang"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

## 2.1 Getting Started

### [2.1.1]{style="colour:grey"} Loading the packages

```{r}
pacman::p_load(ggdist, ggridges, ggthemes,
               colorspace, tidyverse)
```

### [2.1.2]{style="colour:grey"} Data Import

```{r}
exam <- read_csv("Data/Exam_data.csv")
```

## 2.2 Visualising Distribution with Ridgeline Plot

### [2.2.1]{style="colour:grey"} Plotting ridgeline graph: ggridges method

The ridgeline plot below is plotted by using `geom_density_ridges()`.

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    colour = "white"
  ) +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## The code chunk

```{r}
#| eval: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS)) +
  geom_density_ridges(
    scale = 3,
    rel_min_height = 0.01,
    bandwidth = 3.4,
    fill = lighten("#7097BB", .3),
    colour = "white"
  ) +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```
:::

### [2.2.2]{style="colour:grey"} Varying fill colors along the x axis

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```

## The code chunk

```{r}
#| eval: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS,
           fill = stat(x))) +
  geom_density_ridges_gradient(
    scale = 3,
    rel_min_height = 0.01) +
  scale_fill_viridis_c(name = "Temp. [F]",
                       option = "C") +
  scale_x_continuous(
    name = "English Grades",
    expand = c(0, 0)
  ) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  theme_ridges()
```
:::

### [2.2.3]{style="colour:grey"} Mapping the probabilities directly with color

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS,
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail Probability",
                       direction = -1) +
  theme_ridges()
```

## The code chunk

```{r}
#| eval: false
ggplot(exam,
       aes(x = ENGLISH,
           y = CLASS,
           fill = 0.5 - abs(0.5-stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Tail Probability",
                       direction = -1) +
  theme_ridges()
```
:::

## 2.3 Visualising Distribution with Raincloud Plot

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(exam,
       aes(x = RACE,
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left",
            justification = 1.2,
            binwidth = .5,
            dotsize = 2)
```

## The code chunk

```{r}
#| eval: false
ggplot(exam,
       aes(x = RACE,
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left",
            justification = 1.2,
            binwidth = .5,
            dotsize = 2)
```
:::

## 2.4 Finishing touch

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(exam,
       aes(x = RACE,
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left",
            justification = 1.2,
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```

## The code chunk

```{r}
#| eval: false
ggplot(exam,
       aes(x = RACE,
           y = ENGLISH)) +
  stat_halfeye(adjust = 0.5,
               justification = -0.2,
               .width = 0,
               point_colour = NA) +
  geom_boxplot(width = .20,
               outlier.shape = NA) +
  stat_dots(side = "left",
            justification = 1.2,
            binwidth = .5,
            dotsize = 1.5) +
  coord_flip() +
  theme_economist()
```
:::
