---
title: "Hands on Ex03"
---

## **3.1 Learning Outcome**

In this section, I will excerise how to create interactive data visualisation by using functions provided by **ggiraph** and **plotlyr** packages.

## **3.2 Getting Started**

First, write a code chunk to check, install and launch the following R packages:

-   [**ggiraph**](https://davidgohel.github.io/ggiraph/) for making ‘ggplot’ graphics interactive.

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables](https://datatables.net/) that create interactive table on html page.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

-   [**patchwork**](https://patchwork.data-imaginist.com/) for combining multiple ggplot2 graphs into one figure.

The code chunk below will be used to accomplish the task.

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

## **3.3 Importing Data**

```{r}
exam_data <- read_csv("Data/Exam_data.csv",show_col_types = FALSE)
```

## **3.4 Interactive Data Visualisation - ggiraph methods**

Interactive is made with [**ggplot geometries**](https://davidgohel.github.io/ggiraph/reference/#section-interactive-geometries) that can understand three arguments:

-   **Tooltip**: a column of data-sets that contain tooltips to be displayed when the mouse is over elements.

-   **Onclick**: a column of data-sets that contain a JavaScript function to be executed when elements are clicked.

-   **Data_id**: a column of data-sets that contain an id to be associated with elements.

### **3.4.1 Tooltip effect with *tooltip* aesthetic**

Below shows a typical code chunk to plot an interactive statistical graph by using **ggiraph** package. Notice that the code chunk consists of two parts. First, an ggplot object will be created. Next, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) of **ggiraph** will be used to create an interactive svg object.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

Notice that two steps are involved. First, an interactive version of ggplot2 geom (i.e. [`geom_dotplot_interactive()`](https://davidgohel.github.io/ggiraph/reference/geom_dotplot_interactive.html)) will be used to create the basic graph. Then, [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) will be used to generate an svg object to be displayed on an html page.

### **3.5.1 Displaying multiple information on tooltip**

The content of the tooltip can be customised by including a list object as shown in the code chunk below.

```{r}
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

The first three lines of codes in the code chunk create a new field called *tooltip*. At the same time, it populates text in ID and CLASS fields into the newly created field. Next, this newly created field is used as tooltip field as shown in the code of line 7.

### **3.6.1 Customising Tooltip style**

Code chunk below uses [`opts_tooltip()`](https://davidgohel.github.io/ggiraph/reference/opts_tooltip.html) of **ggiraph** to customize tooltip rendering by add css declarations.

```{r}
tooltip_css <- "background-color:pink; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

### **3.6.2 Displaying statistics on tooltip**

Code chunk below shows an advanced way to customise tooltip. In this example, a function is used to compute 90% confident interval of the mean. The derived statistics are then displayed in the tooltip.

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE)) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, linewidth = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

It creates a function called `tooltip` that takes three arguments:

-   `y`: a numeric value representing the mean.

-   `ymax`: a numeric value representing the mean plus the standard error.

-   `accuracy`: a parameter controlling the numeric precision, defaulting to 0.01.

Next, the code creates a ggplot object called `gg_point`. It uses the `exam_data` dataset and sets the `RACE` column as the x-axis. It then adds two layers:

1.  A `stat_summary()` layer using the `GeomInteractiveCol` geom to display the means, and using the `tooltip` function to generate a tooltip for each bar showing the mean maths score and its standard error.

2.  A `stat_summary()` layer using error bars to display the standard errors of the means.

Specifically, the workflow of `stat_summary()` is as follows:

1.  Based on the statistical function specified by the `fun.data` parameter (here, it's `mean_se`, which calculates the mean and standard error), the data is grouped, and the statistics are calculated.

2.  The statistical results for each group are calculated, with the mean stored in `y` and the mean plus the standard error stored in `ymax`.(ggplot2 is performing them behind the scenes based on the `mean_se` specification, and storing the mean in `y` and mean+SE in `ymax`, following its built-in rules.)

3.  At this point, `after_stat()` can access the calculated results `y` and `ymax`.

4.  The function within `after_stat()` (here, it's `tooltip(y, ymax)`) takes the calculated results `y` and `ymax` as inputs and generates a new derived value (in this case, a formatted string based on the mean and standard error).

5.  This derived value is mapped to a particular attribute of the geometric object; in this example, it's the `tooltip` attribute of `GeomInteractiveCol`.

### **3.6.3 Hover effect with *data_id* aesthetic**

Code chunk below shows the second interactive feature of ggiraph, namely `data_id`.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        
```

When `stackgroups = TRUE`, it means that the dots will be stacked or grouped vertically based on their values on the x-axis (which in this case is the "MATHS" variable). Dots with the same x-value will be stacked on top of each other, rather than being spread out horizontally.

If `stackgroups = FALSE` (the default value), the dots would be spread out horizontally, even if they have the same x-value. This can lead to overlapping dots and make it difficult to discern the distribution of data points.

### **3.6.4 Styling hover effect**

In the code chunk below, css codes are used to change the highlighting effect.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

### **3.6.5 Combining tooltip and hover effect**

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

### **3.6.6 Click effect with onclick**

`onclick` argument of ggiraph provides hotlink interactivity on the web.

The code chunk below shown an example of `onclick`.

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```

Interactivity: Web document link with a data object will be displayed on the web browser upon mouse click.

### **3.6.7 Coordinated Multiple Views with ggiraph**

```{r}
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID,
        tooltip= ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

Notice that when a data point of one of the dotplot is selected, the corresponding data point ID on the second data visualisation will be highlighted too.

## **3.7 Interactive Data Visualisation - plotly methods!**

There are two ways to create interactive graph by using plotly, they are:

-   by using *plot_ly()*, and

-   by using *ggplotly()*

```{r}
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)
```

### **3.7.2 Working with visual variable: plot_ly() method**

```{r}
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)
```

### **3.7.3 Creating an interactive scatter plot: ggplotly() method**

```{r}
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)
```

### **3.7.4 Coordinated Multiple Views with plotly**

```{r}
# d <- highlight_key(exam_data)
p1 <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))
```

-   [`highlight_key()`](https://www.rdocumentation.org/packages/plotly/versions/4.9.2/topics/highlight_key) of **plotly** package is used as shared data.

-   two scatterplots will be created by using ggplot2 functions.

-   lastly, [*subplot()*](https://plotly.com/r/subplots/) of **plotly** package is used to place them next to each other side-by-side.

### **3.8.1 Interactive Data Table: DT package**

```{r}
DT::datatable(exam_data, class= "compact")
```

### **3.8.2 Linked brushing: crosstalk method**

```{r}
d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        
```
